---
output: pdf_document
urlcolor: blue
---
# Statistical Inference: Week 4 Project: Part II

Author: Alexander Johannes

Date: 10 August 2021

Link to my Statistical Inference repository: [My Repo](https://github.com/phase-plane/datasciencecoursera/tree/master/Inference)

## Overview

This report details the process of performing some basic statistical analysis on the `ToothGrowth` data set.

## Part II: Basic Inferential Data Analysis

### Step 1: Loading the data and performing some basic analysis:

```{r}
data(ToothGrowth)
head(ToothGrowth)
names(ToothGrowth)
tapply(ToothGrowth$len, ToothGrowth$supp, mean)
```

### Step 2: A basic summary of the data, with a visualization:

```{r, echo=FALSE}
summary(ToothGrowth)
Length <- ToothGrowth$len
Supplement <- ToothGrowth$supp
boxplot(Length~Supplement, data = ToothGrowth, 
        main = "Summary Statistics", col = c("red", "grey"), 
        horizontal = TRUE)
```

### Step 3:  Hypothesis tests (t-test) to compare tooth growth by supp and dose

```{r, echo=FALSE}
Dose05 <- t.test(len ~ supp, 
              data = rbind(ToothGrowth[(ToothGrowth$dose == 0.5) & 
                                         (ToothGrowth$supp == "OJ"),],
                           ToothGrowth[(ToothGrowth$dose == 0.5) & 
                                         (ToothGrowth$supp == "VC"),]), 
              var.equal = FALSE)

Dose1 <- t.test(len ~ supp, 
             data = rbind(ToothGrowth[(ToothGrowth$dose == 1) & 
                                        (ToothGrowth$supp == "OJ"),],
                          ToothGrowth[(ToothGrowth$dose == 1) & 
                                        (ToothGrowth$supp == "VC"),]), 
             var.equal = FALSE)

Dose2 <- t.test(len ~ supp, 
             data = rbind(ToothGrowth[(ToothGrowth$dose == 2) & 
                                        (ToothGrowth$supp == "OJ"),],
                          ToothGrowth[(ToothGrowth$dose == 2) & 
                                        (ToothGrowth$supp == "VC"),]), 
             var.equal = FALSE)


results <- data.frame(
  "p-value" = c(Dose05$p.value, Dose1$p.value, Dose2$p.value),
  "Conf.Low" = c(Dose05$conf.int[1],Dose1$conf.int[1], Dose2$conf.int[1]),
  "Conf.High" = c(Dose05$conf.int[2],Dose1$conf.int[2], Dose2$conf.int[2]),
  row.names = c("Dosage 0.5","Dosage 1","Dosage 2"))

results
```

### Step 4: Conclusions and Assumptions

We reject the null hypothesis for a dosage of .5 and 1, meaning there is a difference in tooth growth by delivery method (at the .05 significance level). 

We have performed our t-test under the assumption of unequal variances.